\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-07-20 paper/paper.tex
Edits to first 5 sections.

\bigskip
\begin{abstract}
% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers tends to be limited; high-reliability consumer products fail infrequently. Due in part to multiple causes of failure, simple parametric models have limited application, being unable to capture relevant lifetime characteristics. On the other hand, more suitable models are difficult to fit because of small sample sizes. If it were available, failure analysis could mitigate identifiability issues; however, consumers do not have this information, nor is it have practical value for them.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout. We fit the GLFP model using a hierarchical Bayesian approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from the full posterior distribution.
\end{abstract}

@@

\newpage
\tableofcontents
\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
Consumers have an interest in accurate assessment of product reliability. Toward this end, there is a need for models that can fit the data together with methods that account for uncertainty.  While computationally convenient, simple parametric models can fail to capture key characteristics found in real data.

% In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout. By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures. By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
% 
% We illustrate this method on hard drive reliability data from a cloud-based computing company.  The data contains both left-truncation and right-censoring.  Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous. 
% 
% \subsection{Motivation}
% 
% Failure distributions of high-reliability consumer products can be difficult to assess.  Even when field data are available, information may be limited for a variety of reasons.  Certain brands may have few products in operation resulting in small sample sizes.  Other products could have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.  In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process. After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability. This highlights the importance of choosing a sufficiently flexible model.

Nonparametric methods, which provide such flexibility are well-known, Kaplan-Meier for example. However, truncation, censoring and finite samples limit the scope of inference for these methods. Parametric models do not have these limitations. Truncated and/or censored data can be accomodated simply by adjusting the likelihood. Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength and incorporate prior information.

Often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality. This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization while accomodating lifetime data with infant mortality and wearout but it can require a lot of data to fit because of this complexity. Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained by  Markov Chain Monte Carlo. \\

% the model captures drive-models with lots of observed failures, while the predicted time to failure for drive-models with less information is over or under estimated.

%\begin{shaded}
  %THIS SHOULD BE SEPARATED OUT
  %At the other extreme, separately fitting the model to each drive-model, is difficult for drive-models where the data are not strongly informative.  Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are chosen.  \\
  %
  %We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  With over 60 drive-models in operation, a hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for each drive-model.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: drive-models with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other drive-models to obtain stable estimates.  This allows us to include more of the hard drive data and more importantly, make valid inferences for each group of hard drive brands.\\
  %
  %
  %Failure data for series systems may be collected at the system level
  %by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
  %model (GLFP), for
  %failure times in a population where some units are susceptible to infant mortality. This model
  %covers a subset of the class of distributions known as bathtub distributions
  %\cite{sujata}. Bathtub distributions are characterized by a U-shaped
  %hazard function, where early failures are due to infant mortality and
  %late failure are due to wearout. While complex systems actually
  %exhibit more than two modes of failure, suitable parametric models
  %which can suitably approximate the overall lifetime distribution are
  %useful. 
  %
  %In the case of complex, high-reliability systems where cause of failure is
  %not available, a model which
  %\begin{enumerate}[a.]
  %\item has parameters that can be estimated from the observed data
  %\item is flexible enough to fit the observed data
  %\item in interpretable with respect to available theory of the underlying process
  %\end{enumerate}
  %is desired. We present a Bayesian, hierarchical modeling approach for
  %grouped failure data. As a benefit of working with Monte Carlo
  %samples, we can easily make inference on a wide range of quantities of
  %interest while accounting for uncertainty. We demonstrate this approach on real data hard
  %drive failure data from a
  %cloud-based storage company.
%\end{shaded}
%
%To model this mixture of failure distributions in a unified framework, Chan and Meeker proposed the Generalized Limited Failure Population (GLFP) model \citet{chan}.  In addition to describing the failure time distribution for products that are susceptible to infant mortality and wearout failure modes, it theorizes an unknown proportion of defective units which are susceptible to defects.    While the GLFP model provides flexibility over a simple or piece-wise parametric model, it can be difficult to fit with limited data  As Chan and Meeker discuss in their original paper, few observed failures in addition to unknown causes of failure cause maximum likelihood estimates to be unstable or impossible to obtain.  Therefore, when multiple populations are of interest, comparisons based on the standard GLFP model are restricted to those with sufficient data and observed failure from each failure mode.  A naive approach to this limitation is to combine the data from all products and fit a single GLFP model.  This model can be well-estimated, but the pooled GLFP model fits poorly.  At the other extreme, fitting a separate GLFP model to each product is problematic for products where the data are weakly informative.  Due to identifiability issues, in these cases the model fit is unstable unless strongly informative priors are placed on the model parameters.  \\
%
%We therefore propose to fit the GLFP model using a hierarchical Bayesian approach.  A hierarchical model is advantageous as allows for borrowing of information to improve estimation of the GLFP parameters for a whole class of products.  Moreover, the hierarchical model is a nice compromise between the aggregate and individual modeling approaches: product brands with lots of failures are estimated precisely; and in cases where little data is available the hierarchical model borrows information from other products to obtain stable estimates.  This allows us to include all of the data and more importantly, make valid inferences for each individual product.
%
%
%Failure data for series systems may be collected at the system level
%by end-users interested only in the lifetime of the system. \citet{chan} proposed the generalized limited failure population
%model (GLFP), for
%failure times in a population where some units are susceptible to infant mortality. This model
%covers a subset of the class of distributions known as bathtub distributions
%\cite{sujata}. Bathtub distributions are characterized by a U-shaped
%hazard function, where early failures are due to infant mortality and
%late failure are due to wearout. While complex systems actually
%exhibit more than two modes of failure, suitable parametric models
%which can suitably approximate the overall lifetime distribution are
%useful. 
%
%In the case of complex, high-reliability systems where cause of failure is
%not available, a model which
%\begin{enumerate}[a.]
%\item has parameters that can be estimated from the observed data
%\item is flexible enough to fit the observed data
%\item in interpretable with respect to available theory of the underlying process
%\end{enumerate}
%is desired. We present a Bayesian, hierarchical modeling approach for
%grouped failure data. As a benefit of working with Monte Carlo
%samples, we can easily make inference on a wide range of quantities of
%interest while accounting for uncertainty. We demonstrate this approach on real data hard
%drive failure data from a
%cloud-based storage company.

@@

An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which an analyst to apply prior information in a straightforward way.  Using the inverse cdf of the SEV distribution,
\begin{equation}
\log t_p = \mu + \sigma \Phi_{SEV}^{-1}(p),
\end{equation}
where $t_p = \argmin\{y:\Phi_{SEV}(\log y)>p\}$. Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.
% The Weibull distribution is widely used in practice as it can model failure distributions with both increasing and decreasing hazard functions \cite{meeker}. The parameter $\sigma$ controls this characteristic: if $\sigma < 1$ the Weibull distribution has an increasing hazard function, $\sigma=1$ corresponds to constant hazard, and if $\sigma > 1$ the hazard is decreasing.

@@

% For the purpose of exposition, we will refer to the probability distributions $F_{d1}$ and $F_{d2}$ as ``failure modes." Specifically, we will refer to $F_{1}$ as the ``early failure mode", and $F_{d2}$ as the ``main failure mode." 
The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

@@

\subsection{Censoring and Truncation}
A common feature of lifetime data is right-censoring. Due to time or budget constraints it is rare all units are observed until failure. If a unit has not yet failed when the data is analyzed it is considered right-censored.  In other words, right-censoring puts a lower bound on the failure time.  
Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit this kind of censoring.

@@

\subsection{Example 1}
We fit the GLFP model on real hard drive failure data from a cloud-based storage company.  Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.  The hard drive data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.  No information is available about the cause of failure.  We first focus on the drive brand model with the most observed failures to demonstrate the GLFP model.  We fit the GLFP using a Bayesian framework and specified the following proper priors to ensure a proper posterior distribution.
$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$

The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.  Given the interpretation of $\pi$ as the proportion of defective units, a range primarily covering the left side of 0.50 seems reasonable and weakly informative.  The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.  Again, this allows for a wide range of hazards, larger than we would expect to see in the data.  The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$. For $t_{0.2,2}$ it is $(3.4,16.6)$. Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.  They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower average time to failure than the wearout mode.  This is helpful for estimation since, as discussed in Chan & Meeker, ML had difficulty identifying the parameters of the GLFP model without making strong assumptions about the data.  \\

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.  The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter estimates for the two Weibull distributions also match intuition.  The wearout failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, an increasing hazard.  The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than for the wearout mode.

@@

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see the Appendix for a more detailed description).

In Figure 1 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.  The solid stepwise curve corresponds to the observed failures.  Censored drives are not plotted.  Standard error bands are calculated using Greenwood's method \cite{green}.  Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.  The non-parametric and parametric model align up quite closely.  Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.  Therefore, a single Weibull model is clearly not flexible enough to model this lifetime distribution.

@@

To borrow strength across different populations, we can
%either share parameters across populations, or 
model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
%For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:

@@

In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters. In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their estimation. For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$. A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration. In section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.

@@

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$). 
\subsection{Prior Distributions}
\label{sec:Prior Distributions}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.

@@

\section{Appendix}
\label{sec:Appendix}
Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  Identifying whether failure data are consistent with a specific family of distributions is difficult to do by examining a histogram or other density estimate.  By properly transforming the axes of a plot of an empirical estimate of fraction failing as a function of time with simultaneous confidence bands provides a graphical test for distributional goodness of fit.

Applying this method to the hard drive data, we start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by
Turnbull, and given in more detail by Meeker and Escobar
\cite{turnbull,meeker}.  For each drive-model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by

>>>>

\bigskip
\begin{abstract}
% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers is limited since high-reliability consumer products fail infrequently. Due in part to multiple causes of failure, simple parametric models have limited application, being unable to capture relevant lifetime characteristics. On the other hand, more suitable models are difficult to fit because of small sample sizes. If failure analysis data were available it could mitigate identifiability issues; however, consumers do not have this information, nor does it have practical value for them.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout. We fit the GLFP model using a hierarchical Bayesian approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from full posterior distributions.
\end{abstract}

@@

\newpage
\tableofcontents
\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
Consumers have an interest in accurate assessment of product reliability. Toward this end, there is a need for models that fit the data together with methods that account for uncertainty.  While computationally convenient, simple lifetime models can fail to capture key characteristics found in real data.

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring early ("infant mortality") due to manufacturing defects.  After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability. This highlights the importance of choosing a sufficiently flexible model.

Nonparametric methods, which provide such flexibility are well-known, Kaplan-Meier for example. However, truncation, censoring and finite samples limit the scope of inference for these methods. Parametric models do not have these limitations. Truncated and/or censored data can be accommodated simply by adjusting the likelihood. Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength across groups, and incorporate prior information.

Often no information is available to identify if a failure is due to infant mortality. This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity.  Bayesian methods offer a means of reducing the amount of data required to estimate the model by leveraging multiple sources of information. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

@@

An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.  Using the inverse cdf of the SEV distribution,
\begin{equation}
\log t_p = \mu + \sigma \Phi_{SEV}^{-1}(p),
\end{equation}
where $t_p = \argmin\{y:\Phi_{SEV}(\log y)>p\}$. Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.

@@

The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

@@

\subsection{Censoring and Truncation}
A common feature of lifetime data is right-censoring. Due to time or budget constraints it is rare all units are observed until failure. If a unit has not yet failed when the data is analyzed it is considered right-censored.  In other words, right-censoring puts a lower bound on the failure time.  Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit right-censoring.

@@

\subsection{Example 1}
We fit the GLFP model on real hard drive failure data from a cloud-based storage company.  Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.  The data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.  No information is available about the cause of failure.  We first focus on the drive brand model with the most observed failures to demonstrate the GLFP model.  We fit the GLFP using a Bayesian framework and specified the following proper priors to ensure a proper posterior distribution.
$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$

The prior on $\pi$ implies with $90\%$ probability $\pi$ is in the interval $(.002, .572)$.  Given the interpretation of $\pi$ as the proportion of defective units, which we would expect to be small, a range primarily covering the left side of 0.50 is reasonable and weakly informative.  The prior on $\sigma_1$ and $\sigma_2$ put the majority of mass on the interval $(0.02, 61.1)$ with $90\%$ probability.  Again, this allows for a wide range of hazards, larger than we would expect to see in the data.  The $t_{0.5,1}$ distribution has a prior $90\%$ interval of $(1.4,14.6)$. For $t_{0.2,2}$ the range is $(3.4,16.6)$. Since these last two parameters are on the scale of log-hours, admitting values from hours to decades, we consider them to be relatively uninformative.  They do, however, help separate the early and late failure modes--the earlier infant mortality mode having a lower mean time to failure than the wearout mode.  This is helpful for estimation since, as discussed in Chan & Meeker, ML had difficulty identifying the two sets of GLFP parameters without making strong assumptions about the data.  \\

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.  The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter estimates for the two Weibull distributions also match intuition.  The wearout failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, an increasing hazard.  The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.

@@

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a more detailed description).

In Figure 1 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.  The solid stepwise curve corresponds to the observed failures.  Censored drives are not plotted.  Standard error bands are calculated using Greenwood's method \cite{green}.  Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.  The non-parametric and parametric model align up quite closely.  Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.  Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution.

@@

To borrow strength across different populations, we can model the population specific parameters hierarchically, allowing the data to inform the hyperparameters.
%For the second option, we model the scales, $\sigma_j$, quantiles, $t_{p_j,d,j}$, and proportions, $\pi_d$, of the component distributions as follows:

@@

In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters. In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their inclusion. For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$. A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration. In section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.

@@

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).

\subsection{Prior Distributions}
\label{sec:Prior Distributions}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.

@@

\section{Appendix}
\label{sec:Appendix}
We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by
Turnbull, and given in more detail by Meeker and Escobar
\cite{turnbull,meeker}.  For each drive-model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by
